---
title: "CobbDouglas"
author: "Brandon Medina"
date: "4/24/2020"
output: html_document
---

## non-linear optimization

This Rmarkdown file demonstrates techniques for non-linear optimization, and is illustrated with data for the Cobb-Douglas production function, $y = A K^\alpha L^\beta$, K capital, L labor, and the exponents the elasticies of production by input.

### Methods demonstrated:
 1. Grid search of space of parameters  
 2. non-linear optimization software, find values of the parameters that minimize the sum of squared errors, errors are (data-predicted)  
 3. non-linear optimizatin software, maximum likelihood- find values of the parameters that maximize the likelihood of the data assuming errors are normally distributed  
 4. log-transform to a linear model  
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in the data

```{r}
CobbDoug = read.csv('CobbDouglasData.csv')
attach(CobbDoug)
```


```{r}
plot(Cap, Q, pch=19, col='red', xlab= 'input values', ylab = 'output',
     main='Cobb-Douglas Production function')
points(Emp, Q, pch=19, col='blue')
grid()
```

A two-dimensional plot would be nice..

```{r}

```

Set up a non-linear loss function

```{r}
CDloss=function(parms, Q, Cap, Emp){
  A = parms[1]
  alpha = parms[2]
  beta = parms[3]
  
  Loss = 0
  for (i in 1:length(Q)){
    Loss = Loss + ((Q[i]-A*(Cap[i]^beta*Emp[i]^alpha))^2)
  }
  
  return (log(Loss))
}

CDmle=function(parms, Q, Cap, Emp){
  A = parms[1]
  alpha = parms[2]
  beta = parms[3]
  
  logLoss = 0
  for (i in 1:length(Q)){
    errorTerm= (Q[i]-A*(Cap[i]^beta*Emp[i]^alpha))
    logLoss = logLoss - dnorm(errorTerm, log=T)
  }
  
  return (logLoss)
}
```

Test the loss function

```{r}
A = 1
alpha = 0.5
beta = 0.5
parms=c(A, alpha, beta)
print(CDloss(parms, Q, Cap, Emp))

# better or worse than
A = 0.5
alpha = 0.5
beta = 0.5
parms=c(A, alpha, beta)
print(CDloss(parms, Q, Cap, Emp))
```
## Grid search
  not the best idea, can get close to an answer 

```{r}

for (A in seq(from=0.01, to = 2, length.out = 4))
  for(a in seq(from=0.01, to = 1,length.out = 6))
   for(b in seq(from=0.01, to = 1, length.out = 6)){
     parms = c(A, a, b)
      print(paste(A, a, b,'        ', CDloss(parms, Q, Cap, Emp)))
  }

```

## non linear optimization

use function optim to find values of A, alpha, beta
find values that minimize the loss function, loss = sum of sqrared errors.  This is a 'least squares' solution

```{r}
##     ?optim
parms = c(2, 0.8, 0.2)
opt = optim(parms, CDloss, Q=Q, Cap=Cap, Emp=Emp)
print(opt)
parms = c(opt$par[1], opt$par[2], opt$par[3])
print(CDloss(parms, Q, Cap,Emp))
```

## maximum likelihood estimation

With the right fit, the residual errors should be independent, normally distributed, with mean 0. likelihood of the parameter set (a, $\alpha$, $\beta$) is

$\prod_0^n \phi(e_i)$

$\phi$ is the standard normal distribution


```{r}
parms = c(2, 0.8, 0.2)
opt = optim(parms, CDmle, Q=Q, Cap=Cap, Emp=Emp)
print(opt)
parms = c(opt$par[1], opt$par[2], opt$par[3])
print(CDmle(parms, Q, Cap,Emp))
print(CDloss(parms,Q,Cap,Emp))
```



## transform to logs

```{r}
logModel = glm(log(Q)~log(Cap)+log(Emp))
summary(logModel)
A = exp(logModel$coefficients[1])
alpha = logModel$coefficients[2]
beta = logModel$coefficients[3]
print(paste('A = ',A,' alpha = ', alpha,' beta =', beta))

# CD loss at these values:
parms=c(A, alpha, beta)
print(paste('CD loss: ', CDloss(parms, Q, Cap, Emp)))
```

$$log(q) = 0.54 + 0.23 * log(k) * 0.77 * log(L)$$

Expo

$$Q = exp(0.54) * K^.23 * L^.77$$

$$A = Exp(.54), \alpha = 0.23, \beta = .77$$

## Questions:
## 1 what are the estimated values of A, alpha and beta?
  $par
[1] 1.3037106 0.8288269 0.2215117

$value
[1] 216230.4

$counts
function gradient 
     178       NA 

$convergence
[1] 0

$message
NULL

[1] 12.97694

## 2 What are the confidence intervals on A, alpha, beta?

A = exp(0.54) = 1.72
  
Confidence interval for $\alpha$: 0.23 $\pm$ 1.96 * 0.03

Confidence interval for $\beta$: 0.77 $\pm$ 1.96 * 0.03

Confidence interval for A: log (A) $\pm$ 1.96 * 0.15
exp(log(A)) = A 

Confidence interval for log(A) is 0.54 $\pm$ 1.96 * 0.15
= (0.25,0.82)

So the 95% confidence interval for A is exp(0.25), exp(0.82) = (1.28, 2.27)

[1] "A =  1.71655367062008  alpha =  0.232343040964735  beta = 0.765310412619359"
[1] "CD loss:  15.7914789583455"

            Estimate Std. Error t value Pr(>|t|)    
(Intercept)  0.54032    0.15151   3.566 0.000663 ***
log(Cap)     0.23234    0.02882   8.061 1.55e-11 ***
log(Emp)     0.76531    0.02735  27.980  < 2e-16 ***

A = exp(.54)
exp(.15)
UCL = exp(.54 + 1.96*.15)
LCL = exp (.54 - 1.96*.15)

UCL CI alpha = 0.23 + 1.96 * 0.03 
LCL CI alpha = 0.23 - 1.96 * 0.03

## 3 is production at constant return to scale? If production is increasing or decreasing return to scale, which input is most expensive

The production would be a decreasing return to scale, with labor being the more expensive input of the variables. 


## 4 What does the values of A mean??

  A is the efficiency factor. fixed value for Alpha and beta. If A is larger, then it decreases the value. 
  
  
  





